MSc Thesis

### 2023-03-15 manuscript/main.tex
catching up commit

Silicon abounds in the earth’s crust, with various silicates, such as silicon dioxide (SiO\textsubscript{2}), comprising about 60\% of the crust by mass. Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as water is absorbed into the plant. Silicates are found in a variety of forms, and vary in their plant availability. Crystalline forms, such as quartz, are highly resistant to weathering, and are poor sources of plant-available silicon, while amorphous forms of SiO\textsubscript{2} are more available. As soils age, an increasing share of the plant-available silicon is derived from biogenic silicates, such as diatom testes or plant phytoliths. Plant phytoliths are amorphous masses of silica, generally similar to opal, that are found throughout the plant body and have highly variable geometries (\cite{piperno_phytoliths_2006}). Phytolith morphologies are generally conserved within taxonomic groups, allowing their use as a tool for paleobotanical investigations. The highly variable morphology among clades, with the conservation of morphology within clades suggests active selection upon the structure of phytoliths. The morphology of phytoliths also varies between organs in the plant. Stem phytoliths tend to be more elongate, with putative structural function, while leaf phytoliths are typically more orbicular, and are likely an adaptation to counter herbivory.

Though widespread benefits of silicon are readily observed in many vascular plant families, the exact mechanism through which silicon acts is still poorly understood (\cite{coskun_controversies_2019}), in part due to the seemingly disparate traits that are promoted under silicon supplementation. When supplemented with silicon, plants are generally more resistant to stress. Silicon supplementation shows efficacy in relieving the negative effects of such abiotic stresses as: soil salinity, soil metal toxicity, cold and heat stress, UV stress, water deficits, and phosphorus deficiencies (\cite{cooke_consistent_2016}). Silicon is also effective at limiting the growth and damage of insect and fungal pests (\cite{fauteux_silicon_2005,massey_herbivore_2007}). Silicon-supplied plants put under stress show transcriptome profiles similar to unstressed plants (\cite{coskun_controversies_2019}). The recently put-forward apoplastic barrier hypothesis suggests that the various benefits gained under silicon supplementation derive from the toughening and sealing action of silicon deposition (\cite{coskun_controversies_2019}). Silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector proteins, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology.The toughness of the depositions also serves a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wear down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}). Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.

@@

One of the most important advances in plant-silicon research was understanding the mechanisms through which silicon is acquired and transported into the plant. Silicon’s most common form in soil solution is silicic acid (H2SiO4), which has a maximum solubility of around 2 mM (\cite{haynes_contemporary_2014}). While there is some evidence that small amounts of silicic acid can be transported during water uptake, this method of transport is insufficient to explain the larger amounts of silicon found in some plant families. Research in rice has identified four proteins that transport silicon into and through the plant body. Two of these proteins (LSi1, LSi2) transport silicic acid from the soil into the roots, while the other two (LSi3, LSi6) act to unload silicic acid from the xylem into leaves and inflorescences (\cite{yamaji_orchestration_2015}). Homologues of these proteins have been identified in other cereal crops, and additional analogous silicon transporter proteins have been discovered in Cucurbitaceae (\cite{reynolds_silicon_2016}). Though not identified, there is a hypothesized fifth protein responsible for loading silicic acid into the xylem (\cite{farooq_silicon_2015}). The expression of these genes, or lack-there-of, can not only influence the total amount of silicon accumulated by the plant, but also its relative distribution, as knockout of LSi6 increases leaf silicon content while decreasing the silicon content of seed husks in rice \cite{yamaji_transporter_2008}. Breeding for silicon content and use-efficiency in crop plants may be crucial to improving crop performance under a changing climate \cite{christian_breeding_2022}. However, we still have a relatively poor understanding surrounding the how genetics influence the silicon phenotype of a plant. Further investigations into how genotypic variation is reflected in the silicon content of plants can aid in the discovery of new genes involved in silicon accumulation, and may provide targets for silicon breeding programs.

@@

Once inside the plant, silicon is deposited in specialized silica cells, forming phytoliths (\cite{waterman_short-term_2021}). Silicon deposits show consistent and taxa-specific morphologies, suggesting evolutionary pressure selecting for these bodies to yield certain functions to the plant (\cite{piperno_phytoliths_2006}). In stems, these phytoliths are often long and narrow, oriented parallel with the shoot, and seem to increase structural rigidity(\cite{stromberg_functions_2016}). The use of silicon as a structural component represents a highly energetically efficient strategy, as silicon is ~10x cheaper on an energy unit basis to produce than lignin (\cite{stromberg_functions_2016}). Stem silicification has been investigated as it relates to lodging resistance in cereal crops, and silicon supplementation has been found to reduce the prevalence of lodging in rice and wheat (\cite{dorairaj_influence_2017,muszynska_mechanistic_2021}). In leaves, phytoliths are typically more stout, though they still increase the mechanical toughness of the leaf (\cite{simpson_still_2017}). This overall toughness, and abundance of phytoliths in leaves likely evolved to limit herbivore damage, rather than improve the growth characteristics as in stem phytoliths (\cite{stromberg_functions_2016}). Interestingly, even in the absence of silicon, plants develop silica cells, and rapidly fill them when silicon becomes available (\cite{waterman_short-term_2021-1}). As phytoliths deposit in the leaves, polymerization is aided by interactions with proteins in the cell wall, which control sites of nucleation (\cite{nawaz_phytolith_2019}). Silicon deposition in the leaves can happen on relatively short time scales, outpacing the accumulation of other defensive compounds such as phenolics (\cite{waterman_short-term_2021}). Thus, silicon-based defences in crop plants may be one of the first lines of active anti-herbivore defence, providing rapid and sensitive responses to herbivory.

@@

Much of today’s plant silicon work is indebted to the pioneering work of Jones and Handreck (\citeyear{jones_silica_1967}) and the subsequent mapping of silicon across the plant kingdom by Takahashi et al. (\citeyear{takahashi_possibility_1990}). Epstein’s seminal \citeyear{epstein_silicon_1999} paper provided a comprehensive review of the state of knowledge in plant silicon, and has spurred a generation of researchers to extend the preliminary findings of the 20th century out across crop production systems and plant ecologies around the world. Silicon is best studied in the grass family (Poaceae) due to the comparatively high silicon content found in most members of the family (often over 1\% of dry weight), as well as the economic importance of domesticated species within the clade (\cite{reynolds_silicon_2016}). Rice, maize, wheat, and barley alone account for one-third of the worlds’ total cultivated land area (\cite{faostat}), and are all domesticated grass species. Silicon supplementation as an agricultural practice has been extensively studied in rice and sugar cane, as these crops tend to deplete soil silicon stocks, necessitating replenishment by application of silicon-rich amendments (\cite{haynes_contemporary_2014,meena_case_2014}). Due to the overall high silicon content of soils globally, Si is rarely truly limiting in soils, though certain forms of silicon are much more plant available than others (\cite{fraysse_surface_2009}). Thus, the applicability and importance of silicon supplementation is unlikely to be realized in more temperate dry-land production systems, particularly in wheat and barley. This does not however nullify the utility of silicon research in these systems, as great work can still be done to improve the manner and efficiency in which these temperate crops utilize the ample silicon available in their soils. Our ability to integrate silicon as a tool for production improvement in dry-land grain production is currently limited by a poor understanding of the genetic controls over silicon accumulation, as well as a limited understanding about the extent to which dry-land cereals utilize silicon in pest-protection.

@@

24 hours after treatment application, I harvested leaf material by clipping all undamaged (i.e. no cricket exposure) leaves. I flash froze leaf tissue liquid nitrogen, freeze dried the tissue, and stored it in a -80ºC freezer. To prepare the leaf powder for XRF analysis, I transferred the dried frozen leaf tissue to 2 mL microcentrifuge tube. I cryoground the leaf tissue at -80ºC in a tissuelyser bead mill for 30 seconds at 60 Hz.

@@

To perfrom the GWAS analysis, I followed the methodology and code published in Gaurav et al. \citedate{gaurav_population_2021}. For brevity, this methodology only describes the steps we took using the data generated from Gaurav et al. \citedate{gaurav_population_2021}. For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript. Prior to the linear regression to assess genetic associations, we first tested our phenotypic data for block effects within sites. For each site, we tested a linear model to see if measured leaf silicon content was significantly associated with block. We tested the model $y \sim Block + Accession$. When we validated that block had a negligible effect ($p>0.1$), we reparameterized leaf silicon measurements at the site level. We converted the raw \%Si to be expressed as the number of standard deviations the sample deviated from the site mean of \%Si. I did this to eliminate a possible confounding effect of inter-site differences that might reduce the power of the GWAS linear regression.

@@

Of the approximately 1700 plants planted, [1300] produced enough leaf material for analysis. Silicon content in \textit{Aegilops tuaschii} ranged from [\textit{value}]\% to [\textit{value}\%]. The various growing environments drove a large amount of variation in silicon content. Overall, my analysis revealed [four] regions of the \textit{Aegilops tauschii} genome that has significant associations with silicon content (Figure \ref{Fig:si_peak_plot}). One of these genomes was on chromosome 4S, near a known gene analogue to \textit{Lsi1}, a silicon transporter protein. My results for manganese content are less clear. I detected no regions that met the threshold for significance, though there were three that had pronounced peaks relative to the average response (Figure \ref{Fig:mn_peak_plot}). Within the plants, silicon and manganese content were correlated ($R^2$ = 0.15, p = 0.049) (Figure 3).

>>>>

Silicon abounds in the earth’s crust, with various silicates, such as silicon dioxide (SiO\textsubscript{2}), comprising about 60\% of the crust by mass. Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as water is absorbed into the plant. Silicates are found in a variety of forms, and vary in their plant availability. Crystalline forms, such as quartz, are highly resistant to weathering, and are poor sources of plant-available silicon, while amorphous forms of SiO\textsubscript{2} are more available \cite{fraysse_surface_2009}. As soils age, an increasing share of the plant-available silicon is derived from biogenic silicates, such as diatom testes or plant phytoliths \cite{de_tombeur_plants_2020}. Uptake and usage of silicon is not uniform throughout the vascular plants. Plant clades vary in their expression of silicon transporter proteins and in the relative abundance of phyotliths in their tissues. Plant phytoliths are amorphous masses of silica, chemically similar to opal, that are found throughout the plant body and have highly variable geometries (\cite{piperno_phytoliths_2006}). Phytolith morphologies are generally conserved within taxonomic groups, allowing their use as a tool for paleobotanical investigations. The highly variable morphology among clades, with the conservation of morphology within clades suggests active selection upon the structure of phytoliths. The morphology of phytoliths also varies between organs in the plant. Stem phytoliths tend to be more elongate, with putative structural function, while leaf phytoliths are typically more orbicular, and are likely an adaptation to counter herbivory.

Silicon deposition may create apoplastic barriers that seal and toughen the plant tissue, which may benefit the plant in a multitude of ways (\cite{coskun_controversies_2019}). Under this hypothesis, silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector compounds, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology (\cite{coskun_controversies_2019}). Empirical evidence shows that silicon is effective at limiting the growth and damage of insect and fungal pests (\cite{fauteux_silicon_2005,massey_herbivore_2007}). The toughness of the depositions also serves a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wearing down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}). Indeed, when supplemented with silicon plants are generally more resistant to a wide spectrum of stressors, including soil salinity, soil metal toxicity, cold and heat stress, UV stress, water deficits, and phosphorus deficiencies (\cite{cooke_consistent_2016}). Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.

%Though widespread benefits of silicon are readily observed in many vascular plant families, the exact mechanism through which silicon acts is still poorly understood (\cite{coskun_controversies_2019}), in part due to the seemingly disparate traits that are promoted under silicon supplementation. When supplemented with silicon, plants are generally more resistant to stress. Silicon supplementation shows efficacy in relieving the negative effects of such abiotic stresses as: soil salinity, soil metal toxicity, cold and heat stress, UV stress, water deficits, and phosphorus deficiencies (\cite{cooke_consistent_2016}). Silicon is also effective at limiting the growth and damage of insect and fungal pests (\cite{fauteux_silicon_2005,massey_herbivore_2007}). Silicon-supplied plants put under stress show transcriptome profiles similar to unstressed plants (\cite{coskun_controversies_2019}). The recently put-forward apoplastic barrier hypothesis suggests that the benefits gained under silicon supplementation derive from the toughening and sealing action of silicon deposition on the apoplast (\cite{coskun_controversies_2019}). Silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector compounds, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology. The toughness of the depositions also serves a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wearing down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}). Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.

@@

One of the most important advances in plant-silicon research was understanding the mechanisms through which silicon is acquired and transported into the plant. Silicon’s most common form in soil solution is silicic acid (H2SiO4), which has a maximum solubility of around 2 mM (\cite{haynes_contemporary_2014}). While there is some evidence that small amounts of silicic acid can be transported during water uptake, this method of transport is insufficient to explain the larger amounts of silicon found in some plant families. Research in rice has identified four gene products that transport silicon into and through the plant body. Two of these (LSi1, LSi2) transport silicic acid from the soil into the roots, while the other two (LSi3, LSi6) act to unload silicic acid from the xylem into leaves and inflorescences (\cite{yamaji_orchestration_2015}). Orthologs of these proteins have been identified in other cereal crops, and additional analogous silicon transporter proteins have been discovered in the Cucurbitaceae (\cite{reynolds_silicon_2016}). Though not identified, there is a hypothesized fifth protein responsible for loading silicic acid into the xylem (\cite{farooq_silicon_2015}). The expression of these genes, or lack-there-of, can not only influence the total amount of silicon accumulated by the plant, but also its relative distribution, as knockout of LSi6 increases leaf silicon content while decreasing the silicon content of seed husks in rice \cite{yamaji_transporter_2008}. Breeding for silicon content and use-efficiency in crop plants may be crucial to improving crop performance under a changing climate \cite{christian_breeding_2022}. However, we still have a relatively poor understanding surrounding the how genetics influence the silicon phenotype of a plant. Further investigations into how genotypic variation is reflected in the silicon content of plants can aid in the discovery of new genes involved in silicon accumulation, and may provide targets for silicon breeding programs.

@@

Once inside the plant, silicon is deposited in specialized silica cells, forming phytoliths (\cite{waterman_short-term_2021}). Silicon deposits show consistent and taxa-specific morphologies, suggesting evolutionary pressure selecting for these structures to yield certain functions to the plant (\cite{piperno_phytoliths_2006}). In stems, these phytoliths are often long and narrow, oriented parallel with the shoot, and seem to increase structural rigidity(\cite{stromberg_functions_2016}). The use of silicon as a structural component represents a highly energetically efficient strategy, as silicon is ~10x cheaper on an energy unit basis to produce than lignin (\cite{stromberg_functions_2016}). Stem silicification has been investigated as it relates to lodging resistance in cereal crops, and silicon supplementation has been found to reduce the prevalence of lodging in rice and wheat (\cite{dorairaj_influence_2017,muszynska_mechanistic_2021}). In leaves, phytoliths are typically more stout, though they still increase the mechanical toughness of the leaf (\cite{simpson_still_2017}). This overall toughness, and abundance of phytoliths in leaves likely evolved to limit herbivore damage, rather than improve the growth characteristics as in stem phytoliths (\cite{stromberg_functions_2016}). Interestingly, even in the absence of silicon, plants develop silica cells, and rapidly fill them when silicon becomes available (\cite{waterman_short-term_2021-1}). As phytoliths develop in the leaves, polymerization of monosilicic acid is aided by interactions with proteins in the cell wall, which act as sites of nucleation (\cite{nawaz_phytolith_2019}). Silicon deposition in the leaves can happen on relatively short time scales, outpacing the accumulation of other defensive compounds such as phenolics (\cite{waterman_short-term_2021}). Thus, silicon-based defences in crop plants may be one of the first lines of active anti-herbivore defence, providing rapid and sensitive responses to herbivory.

@@

Much of today’s plant silicon work is indebted to the pioneering work of Jones and Handreck (\citeyear{jones_silica_1967}) and the subsequent mapping of silicon across the plant kingdom by Takahashi et al. (\citeyear{takahashi_possibility_1990}). Epstein’s seminal \citeyear{epstein_silicon_1999} paper provided a comprehensive review of the state of knowledge in plant silicon, and has spurred a generation of researchers to extend the preliminary findings of the 20th century out across crop production systems and plant ecologies around the world. Silicon is best studied in the grass family (Poaceae) due to the comparatively high silicon content found in most members of the family (often over 1\% of dry weight), as well as the economic importance of domesticated species within the clade (\cite{reynolds_silicon_2016}). Rice, maize, wheat, and barley alone account for one-third of the worlds’ total cultivated land area (\cite{faostat}), and are all domesticated grass species. Silicon supplementation as an agricultural practice has been extensively studied in rice and sugar cane, as these crops tend to deplete soil silicon stocks, necessitating replenishment by application of silicon-rich amendments (\cite{haynes_contemporary_2014,meena_case_2014}). Due to the overall high silicon content of soils globally, Si is rarely truly limiting in soils, though certain forms of silicon are much more plant available than others (\cite{fraysse_surface_2009}). Thus, the applicability and importance of silicon supplementation is unlikely to be realized in more temperate dry-land production systems, particularly in wheat and barley. This does not however nullify the utility of silicon research in these systems, as great work can still be done to improve the manner and efficiency in which these temperate crops utilize the ample silicon available in their soils. Our ability to integrate silicon as a tool for production improvement in crop production is currently limited by a poor understanding of the genetic controls over silicon accumulation, as well as a limited understanding about the extent to which crops utilize silicon in pest-protection.

@@

18 hours after treatment application, I harvested leaf material by clipping all undamaged (i.e. no cricket exposure) leaves. I harvested three fully expanded upper leaves the plants, and split the plants in half (COME BACK HERE!!) I flash froze leaf tissue in liquid nitrogen, freeze dried the tissue, and stored it in a -80ºC freezer. To prepare the leaf powder for XRF analysis, I transferred the dried frozen leaf tissue to 2 mL microcentrifuge tube. I cryoground the leaf tissue at -80ºC in a tissuelyser bead mill for 30 seconds at 60 Hz.

@@

To perfrom the GWAS analysis, I followed the methodology and code published in Gaurav et al. \citedate{gaurav_population_2021}. For brevity, this methodology only describes the steps I took using the data generated from Gaurav et al. \citedate{gaurav_population_2021}. For full details on how they generated the sequence data and prepared the final data sets refer to their manuscript. Prior to the linear regression to assess genetic associations, I first tested my phenotypic data for block effects within sites. For each site, we tested a linear model to see if measured leaf silicon content was significantly associated with block. I tested the model $Si \sim Block + Accession$. When I validated that block had a non-significant effect ($p>0.1$), I reparameterized leaf silicon measurements at the site level; I converted the raw \%Si to be expressed as the number of standard deviations the sample deviated from the site mean of \%Si. I did this to eliminate a possible confounding effect of inter-site differences that might reduce the power of the GWAS linear regression.

@@

Of the approximately 1700 plants planted, [1300] produced enough leaf material for analysis. Silicon content in \textit{Aegilops tuaschii} ranged from [\textit{value}]\% to [\textit{value}\%]. Site accounted for a large amount of variation in silicon content, suggesting a strong environmental effect on the silicon phenotype. Overall, my analysis revealed [four] regions of the \textit{Aegilops tauschii} genome that has significant associations with silicon content (Figure \ref{Fig:si_peak_plot}). One of these genomes was on chromosome 4S, near a known gene analogue to \textit{Lsi1}, a silicon transporter protein. My results for manganese content are less clear. I detected no regions that met the threshold for significance, though there were three that had pronounced peaks relative to the average response (Figure \ref{Fig:mn_peak_plot}). Within the plants, silicon and manganese content were correlated ($R^2$ = 0.15, p = 0.049) (Figure 3).
